---
title: 'Graduates'
author: "Kimutai"
date: "6/2/2021"
output: 
  bookdown::html_document2:
    css: grad_stylesheets.css
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F, 
                      out.width = "95%")
```

# This data is based off the article from 538 data [Guide to picking a college major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/).

```{r}

library("tidyverse")
library("scales")

#Loading a custom theme.
theme_set(theme(panel.background = element_rect("#cef9f7"),
                plot.background =  element_rect("#e7f9f8"),
                panel.grid.major.y = element_line("#9291b8"),
                legend.background = element_rect("#e7f9f8"),
                legend.position = "bottom" ))
#Reading in the data set.
graduates <- read_csv("recent_grads.csv")
```
## The top earning major earn within a range of the following.

STEM degree majors tend to earn more than their humanities and arts counterparts.  
However the humanities and arts are more populous than STEM counterparts.  
A college degree is no guarantee of economic success but the choice of a major can boost the odds.  
Top earning majors are as follows.




```{r}
graduates %>% 
  arrange(desc(median)) %>% 
  select(major_category, median, major, p25th, p75th) %>% 
  head(20) %>% 
  mutate(major = str_to_title(major), 
         major = fct_reorder(major, median)) %>% 
  ggplot(aes(x = median, y = major, col = major_category)) +
  geom_point() +
  geom_errorbar(aes(xmin = p25th, xmax = p75th)) +
  expand_limits(x = 0) +
  labs(title = "Earning ranges of top earning degrees",
       x = "Earnings ranges", y = NULL) + 
  scale_x_continuous(labels = dollar) +
  guides(color = guide_legend(nrow = 2))
```

## A breakdown of the most populous degree in terms of the gender totals. 



```{r}
graduates %>% 
  filter(!is.na(total)) %>% 
  arrange(desc(total)) %>% 
  head(20) %>% 
  mutate(major = fct_reorder(major, total)) %>% 
  pivot_longer(men:women, 
               names_to = "gender", 
               values_to = "number")%>% 
  ggplot(aes(x = major,y = number, fill = gender)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = comma) +
  expand_limits(y = 0, x = 0) +
  coord_flip() +
  labs(x = NULL)
  
```



## Major categories breakdown in terms of the gender.


```{r}
graduates %>% 
  filter(!is.na(total)) %>% 
    pivot_longer(cols = c(men, women), 
                           names_to = "gender", 
                           values_to = "number") %>% 
  group_by(major_category, gender) %>% 
  summarize(n = sum(number)) %>% 
  filter(!is.na(n)) %>% 
  ggplot(aes(x = fct_reorder(major_category, n),
             y = n, fill = gender)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme(panel.background = element_rect("#cef9f7"),
        plot.background =  element_rect("#e7f9f8"),
        panel.grid.major.y = element_line("#9291b8")) +
  scale_y_continuous(labels = comma) +
  #scale_fill_manual(values = c(women = "#E54F5F",men = "#214490")) +
  labs(title = "Men & Women in Various Disciplines",
       x = NULL,
       y = "No. of people")
       
```












